Death By Dropdown?

### 2025-04-02 conference-materials/2025_shiny-conf/slides.qmd
Refines reactable and echarts4r examples in slides
- Updates code snippets for clarity and BID principles.
- Simplifies code logic, enhances comments, and integrates quarto support.
- Adds code highlighting functionality to snippets for enhanced visual comprehension and user experience.

```r
reactable(
  soundpulse_data,
  # Stage 1: Cognitive Load Theory - Grouped structure reduces complexity
  groupBy = "Claim",

  # Stage 2: Processing Fluency - Color coding for instant insights
  columns = list(
    Metric1 = colDef(
      style = function(value) {
        # Color coding enhances processing fluency
        if (value > 50) {
          list(
            background = "rgba(0, 128, 128, 0.2)", 
            color = "#006666",
            fontWeight = "bold"
          )
        } else if (value < 30) {
          list(
            background = "rgba(255, 140, 0, 0.2)", 
            color = "#dd8500",
            fontWeight = "bold"
          )
        }
      }
    )
  )
)
```

@@

```r
reactable(
  soundpulse_data,
  # Stage 3: Principle of Proximity - Group related elements visually
  columnGroups = list(
    colGroup(
      name = "System 1 - Quick Insights", 
      columns = c("Metric1", "Metric2")
    ),
    colGroup(
      name = "System 2 - Detailed Analysis", 
      columns = c("Metric3", "Metric4")
    )
  ),

  # Stage 2 & 3: Enhanced Processing Fluency with bar charts
  columns = list(
    Metric1 = colDef(
      cell = function(value) create_bar_chart_cell(value, "metric1"),
      html = TRUE
    )
  ),

  # Progressive disclosure supports Cognitive Load Theory
  details = function(index) {
    # Only show details when requested
    detailed_data <- segment_details[
      segment_details$claim_id == soundpulse_data$Claim[index], 
    ]
    div(
      style = "padding: 1rem",
      reactable(detailed_data, outlined = TRUE, striped = TRUE)
    )
  }
)
```

@@

```r
# Create a static gauge chart
create_static_gauge <- function(value, column_name, tooltip_text = NULL) {
  chart <- e_charts() |>
    e_gauge(
      value,
      startAngle = 180,
      endAngle = 0,
      detail = list(formatter = "{value}%", fontSize = 14),
      pointer = list(show = FALSE),
      axisLine = list(
        lineStyle = list(
          color = list(
            c(round(value / 100, 2), "#1c6d7d"),
            c(1, "#d2d2d2")
          ),
          width = 50
        )
      )
    ) |>
    e_title(column_name)

  if (!is.null(tooltip_text)) {
    chart <- chart |> e_tooltip(
      formatter = htmlwidgets::JS(
        sprintf("function(params) { return '%s'; }", tooltip_text)
      )
    )
  }

  return(chart)
}
```

@@

```r
# Create a static bar chart with benchmark
create_static_bar <- function(
  value, column_name, benchmark = NULL, 
  tooltip_text = NULL, frame_type = "progress") {
  # Chart creation code...

  # Add benchmark line if provided
  if (!is.null(benchmark)) {
    chart <- chart |>
      e_mark_line(
        data = list(xAxis = benchmark, name = "Benchmark"),
        symbol = "none",
        lineStyle = list(color = "#ff9000", width = 2, type = "dashed"),
        label = list(show = TRUE, formatter = "{b}: {c}%")
      )
  }

  return(chart)
}

# Usage in layout
layout_columns(
  col_widths = c(4, 4, 4),
  div(create_static_bar(65, "KPI 1", 50, "Benchmark is industry average")),
  div(create_static_bar(45, "KPI 2", 50, "Benchmark is industry average")),
  div(create_static_bar(72, "KPI 3", 50, "Benchmark is industry average"))
)
```

@@

```r
# Frame toggle UI
frame_toggle <- div(
  class = "frame-toggle-container",
  span(class = "frame-toggle-label", "Framing:"),
  div(
    class = "btn-group",
    role = "group",
    `aria-label` = "Framing toggle",
    tags$button(
      id = "progress-btn",
      class = "btn btn-sm btn-primary active",
      type = "button",
      "Progress"
    ),
    tags$button(
      id = "gap-btn",
      class = "btn btn-sm btn-outline-primary",
      type = "button",
      "Gap"
    )
  )
)

# JavaScript for toggling
tags$script(HTML("
  function toggleFraming(framingType) {
    if (framingType === 'progress') {
      // Show progress charts, hide gap charts
      document.querySelectorAll('.framing-chart:nth-child(1)').forEach(function(chart) {
        chart.classList.remove('hidden-chart');
      });
      document.querySelectorAll('.framing-chart:nth-child(2)').forEach(function(chart) {
        chart.classList.add('hidden-chart');
      });
    } else {
      // Show gap charts, hide progress charts
      document.querySelectorAll('.framing-chart:nth-child(1)').forEach(function(chart) {
        chart.classList.add('hidden-chart');
      });
      document.querySelectorAll('.framing-chart:nth-child(2)').forEach(function(chart) {
        chart.classList.remove('hidden-chart');
      });
    }
  }
"))
```

>>>>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "4-5|6-18|19-23"

# Packages: reactable, dplyr, bslib
reactable(
  soundpulse_data,
  # Stage 1: Reduce cognitive load with grouped structure
  groupBy = "Song",
  # Stage 2: Enhance processing fluency with color coding
  columns = list(
    `Metric 1` = colDef(
      style = function(value) {
        # Color coding for instant comprehension
        case_when(
          value > 60 ~ list(background = "#1c6d7d", color = "white"),
          value > 40 ~ list(background = "#35a4ae", color = "white"),
          TRUE ~ list(background = "#ffca90", color = "black")
        )
      }
    )
  ),
  # Stage 1: Progressive disclosure reduces overload
  details = function(index) {
    song_details <- filter(details_data, song_id == soundpulse_data$Song[index])
    card(reactable(song_details, outlined = TRUE))
  }
)
```

@@

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "4-14|15-49"

# Packages: reactable, dplyr, bslib
reactable(
  soundpulse_data,
  # Stage 3: Group related elements (Principle of Proximity)
  columnGroups = list(
    colGroup(
      name = "Core Metrics",
      columns = c("Metric 1", "Metric 2", "Metric 3", "Metric 4")
    ),
    colGroup(
      name = "Key Performance",
      columns = c("KPI 1", "KPI 2", "KPI 3")
    )
  ),
  # Stage 1: Dynamic legend reduces choice complexity (Hick's Law)
  defaultColDef = colDef(
    header = function(value) {
      layout_inline(
        value,
        span(
          style = css(
            display = "inline-block",
            margin_left = "5px",
            height = "10px", 
            width = "10px",
            border_radius = "50%",
            background_color = "#1c6d7d"
          )
        )
      )
    }
  ),
  # Combine multiple visual elements for age groups
  columns = list(
    Age = colDef(
      cell = function(value) {
        colors <- c(
          "18-24" = "#1c6d7d",
          "25-34" = "#35a4ae", 
          "35-44" = "#98d3ca",
          "45-54" = "#ffca90",
          "55-64" = "#1c6d7d",
          "65-74" = "#dd8500"
          )

        span(value, style = css(color = colors[value], font_weight = "bold"))
      }
    )
  )
)
```

@@

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-30"

# Packages: echarts4r, bslib
# Create intuitive gauge chart (Stage 2: Processing Fluency)
create_gauge_chart <- function(value, title) {
  e_charts() |>
    e_title(title) |>
    e_gauge(
      value,
      startAngle = 180,
      endAngle = 0,
      detail = list(formatter = "{value}%"),
      axisLine = list(
        lineStyle = list(
          color = list(
            c(value/100, "#1c6d7d"),  # Filled portion
            c(1, "#e9e9e9")           # Empty portion
          ),
          width = 30
        )
      )
    )
}

# Usage in dashboard (or use purrr::map)
layout_column_wrap(
  width = 1 / 4,
  create_gauge_chart(mean(music_data$`Metric 1`), "Metric 1"),
  create_gauge_chart(mean(music_data$`Metric 2`), "Metric 2"),
  create_gauge_chart(mean(music_data$`Metric 3`), "Metric 3"),
  create_gauge_chart(mean(music_data$`Metric 4`), "Metric 4")
)
```

@@

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-29"

# Packages: echarts4r, bslib
# Create benchmark bar (Stage 4: Anchoring Effect)
create_benchmark_bar <- function(value, title, benchmark = 50) {
  e_charts() |>
    e_title(title) |>
    e_bar(
      value,
      legend = list(show = FALSE),
      showBackground = TRUE,
    ) |>
    e_flip_coords() |>
    e_labels(position = "inside") |>
    e_y_axis(show = FALSE) |>
    e_x_axis(show = FALSE, min = 0, max = 100) |>
    e_hide_grid_lines() |>
    e_mark_line(
      data = list(yAxis = benchmark),
      lineStyle = list(color = "#dd8500", type = "dashed", width = 2),
      label = list(show = TRUE, formatter = "Benchmark")
    )
}

# Usage in dashboard (or use purrr::map)
layout_columns(
  width = 1 / 3,
  card(create_benchmark_bar(mean(music_data$popularity), "Popularity")),
  card(create_benchmark_bar(mean(music_data$danceability), "Danceability")),
  card(create_benchmark_bar(mean(music_data$energy), "Energy"))
)
```

@@

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-3|6-12|14-20|23-24|25-40|42-43|45-59"

# Packages: shiny, echarts4r, dplyr, purrr, bslib
# UI with framing toggle (Stage 4: Framing & Loss Aversion)
ui <- fluidPage(
  titlePanel("SoundPulse Dashboard"),

  # Stage 4: Framing toggle
  radioButtons(
    "framing", "Framing:", 
    choices = c("Progress" = "progress", "Gap" = "gap"),
    selected = "progress",
    inline = TRUE
  ),

  # Layout for charts using programmatically generated outputs
  layout_column_wrap(
    width = 1 / 3,
    card(echarts4rOutput("chart_popularity")),
    card(echarts4rOutput("chart_danceability")),
    card(echarts4rOutput("chart_energy"))
  )
)

# Server logic
server <- function(input, output, session) {
  # Reactive expression for framed data
  framed_data <- reactive({
    soundpulse_data |>
      mutate(
        display_value = if_else(
          input$framing == "progress",
          value,
          abs(benchmark - value)
        ),
        title = if_else(
          input$framing == "progress",
          metric,
          paste0(metric, " Gap")
        )
      )
  })

  # Grab KPI column names
  kpi_col_names <- c("popularity", "danceability", "energy")

  # Generate all chart outputs dynamically
  walk(kpi_col_names, function(kpi_col) {
    output_id <- paste0("chart_", kpi_col)

    output[[output_id]] <- renderEcharts4r({
      metric_data <- filter(framed_data(), metric == kpi_col)

      # Function from previous slide
      create_benchmark_bar(
        metric_data$display_value,
        metric_data$title,
        metric_data$benchmark
      )
    })
  })
}
```
